Omics-based Single Step Trait Prediction

### 2017-09-14 paper/manuscript.tex
paper, explain partitions of W matrix

\Genetics2level{Prediction}
\subsection{Kernels}
BLUP models are computationally attractive when the number of features $p$
exceeds the number of genotypes $n$ and are equivalent to a selection index
when fixed effects have been accomodated by the dependent variable 
\cite{Mrode2014}.
Depending on the data set, up to three predictors were available for agronomic
trait predictions, namely pedigree data (P), genomic data (G) and
transcriptompic data (T).
The corresponding feature matrix of the inbred lines ($\mathbf{M}$) has
dimensions $n \times p$ where $n$ pertains to the number of genotypes.
Note that, for the sake of clarity, we do not distinguish between models for
Exp1 and Exp2 at this point.
In the case of Exp2 the term `genotype' refers strictly to the parent lines.
In generating kernels for each predictor, all features in $\mathbf{M}$ were
centered and standardized to unit variance.
Then, each kernel can be defined as

@@

\subsection{Single step prediction}
If the genetic values are fully captured by the features in $\mathbf{M}$,
Eq.~\ref{eq:KBLUPModel} can then be expressed as

@@

\cite{Fernando2014} because the ridge regression model is equivalent to the
BLUP model~\cite{Ruppert2003}.
In Eq.~\ref{eq:KBLUPModelSimplified}, $\mathbf{M}$ is a matrix of feature
covariates and $\boldsymbol{\alpha}$ is the vector of random, partial-regression
coefficients of the feature covariates.

@@

where $\boldsymbol{\hat{\alpha}}$ is the solution to the ridge regression model
(Eq.~\ref{eq:KBLUPModel}).

Consider now the situation in which one predictor is complete in the sense that
the whole population is equipped with features for that predictor whereas only
a subset of that population has information on a second predictor.
We can take a similar route as in \citeNP{Fernando2014} and impute covariates 
of the incomplete predictor by using covariates of the complete predictor.
Let the subscript $1$ denote individuals covered only by the complete
predictor whereas individuals covered by both, the complete and the
incomplete predictor, are indicated by the subscript $2$.
Further, we center the covariates of the incomplete predictor and denote it as
$\mathbf{W}$.
Correspondingly, $\mathbf{K}$ corresponds to the kernel for the complete
predictor.
In this case we need to partition the vector of genetic values $\mathbf{g}$
into a component $\mathbf{g}_{1}$ (\textit{i.e.,} genetic values of genotypes
covered only by the complete predictor) and a component
$\mathbf{g}_{2} = \mathbf{W}_{2}\boldsymbol{\alpha}$ (\textit{i.e.,} genetic
values of genotypes covered by both predictors, where $\boldsymbol{\alpha}$
pertains to the random feature effects of the incomplete predictor).
Under multivariate normality, the vector $\mathbf{g}_1$ can be written as the
sum of the conditional expectation given $\mathbf{g}_2$ and a residual error
term:
\begin{align} \label{eq:mrna1}
	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
	&= \mathbf{K}_{12}\mathbf{K}^{-1}_{22}\mathbf{W}_{2}
	\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{K}_{12}
	\mathbf{K}^{-1}_{2}
	\boldsymbol{\hat{\alpha}}) \\
	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon},
\end{align}

where $\mathbf{K}_{ij}$ are partitions of $\mathbf{K}$ that correspond to 
$\mathbf{g}_{1}$ and $\mathbf{g}_{2}$, respectively~\cite{Fernando2014} and
where $\mathbf{\hat{g}}_{1}$ pertains to the imputed genetic values in the
incomplete predictor.

>>>>

\Genetics2level{Prediction}
\subsection{Kernels}
BLUP models are computationally attractive when the number of features $p$
exceeds the number of genotypes $n$ and are equivalent to a selection index
when fixed effects have been accomodated by the dependent variable 
\cite{Mrode2014}.
Depending on the data set, up to three predictors were available for agronomic
trait predictions, namely pedigree data (P), genomic data (G) and
transcriptompic data (T).
The corresponding feature matrix of the inbred lines ($\mathbf{M}$) has
dimensions $n \times p$ where $n$ pertains to the number of genotypes.
For the sake of clarity, we do not distinguish between models for Exp1 and Exp2
at this point.
It should be noted, however, that in the case of Exp2 the term `genotype' refers
strictly to the parent lines.
For the complete predictor we generated kernels by centering and standardizing
all features in $\mathbf{M}$ to unit variance:

@@

\subsection{Single step prediction}
If the genetic values are fully captured by the features in $\mathbf{M}$,
Eq.~\ref{eq:KBLUPModel} can be expressed as

@@

\cite{Fernando2014} because the ridge regression model is equivalent to the
BLUP model~\cite{Ruppert2003}.
In the previous equation, $\mathbf{M}$ is a matrix of feature covariates and
$\boldsymbol{\alpha}$ is the vector of random, partial-regression coefficients
of the feature covariates.

@@

where $\boldsymbol{\hat{\alpha}}$ is the solution to the ridge regression model.

Consider now the situation in which one predictor is complete in the sense that
the whole population is equipped with features for that predictor whereas only
a subset of that population has information on a second predictor.
We can take a similar route as in \citeNP{Fernando2014} and impute covariates 
of the incomplete predictor by using covariates of the complete predictor.
Let the subscript $1$ denote individuals covered only by the complete
predictor whereas individuals covered by both, the complete and the
incomplete predictor, are indicated by the subscript $2$.
Let us further denote the matrix of centered covariates of the incomplete
predictor as
\begin{equation} \label{eq:incomplete-covariates}
\mathbf{W} =
\begin{bmatrix}
  \mathbf{\hat{W}}_{1} \\
  \mathbf{W}_{2}
 \end{bmatrix},
\end{equation}
where observed feature covariates are stored in $\mathbf{W}_{2}$ while marker
covariates for genotypes not covered by the incomplete predictor
($\mathbf{\hat{W}}_{1}$) need to be imputed.
In this case we need to partition the vector of genetic values $\mathbf{g}$
into a component $\mathbf{g}_{1}$ (\textit{i.e.,} genetic values of genotypes
covered only by the complete predictor) and a component
$\mathbf{g}_{2} = \mathbf{W}_{2}\boldsymbol{\alpha}$ (\textit{i.e.,} genetic
values of genotypes covered by both predictors where $\boldsymbol{\alpha}$
pertains to the random feature effects of the incomplete predictor).
Under multivariate normality, the vector $\mathbf{g}_1$ can be written as the
sum of the conditional expectation given $\mathbf{g}_2$ and a residual error
term:
\begin{align} \label{eq:mrna1}
	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
	&= \mathbf{K}_{12}\mathbf{K}^{-1}_{22}\mathbf{W}_{2}
	\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{K}_{12}
	\mathbf{K}^{-1}_{2}
	\boldsymbol{\hat{\alpha}}) \\
	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon},
\end{align}

where $\mathbf{K}_{ij}$ are partitions of $\mathbf{K}$ that correspond to 
$\mathbf{g}_{1}$ and $\mathbf{g}_{2}$, respectively~\cite{Fernando2014} and
where $\mathbf{\hat{g}}_{1}$ pertains to the imputed genetic values of genotypes
not equipped with the incomplete predictor.
